Minimizing post shock forecasting error using disparate information

### 2020-06-02 manuscript/manuscript.tex
updated example; updated bootstrap; fixed some references; still need to do more

We will suppose that an analyst has time series data ($y_{i,t}$,$\x_{i,t}$), 
$t = 1$, $\ldots$, $T_i$, $i = 1$, $\ldots$, $n+1$, where $y_{i,t}$ is a 
scalar response and $\x_{i,t}$ is a vector of covariates that are revealed to 
the analyst prior to the observation of $y_{1,t}$.  Suppose that the analyst 
is interested in forecasting $y_{1,t}$, the first time series in the 
collection.
%$y_1, y_2, \ldots$.  
%Given each time point $t \geq 1$, let $\x_{1,t}$ be a vector of covariates 
%revealed prior to the observation of $y_{1,t}$.  
To gauge the performance of a procedure that produces forecasts 
$\{\hat y_{1,t}, t= 1,2,\ldots\}$ given time horizon $T_1$, we consider the 
average forecast risk
$$
  R_T = \frac{1}{T}\sum_{t=1}^T\E{\hat y_{1,t} - y_{1,t}}^2
$$
in our analyses. In this article, we consider a dynamic panel data 
model with autoregressive structure similar to that in \citet{blundell1998initial}. 
%without past time period covariate 
%information and with a single present time covariate.  
Our dynamic panel model includes an additional shock effect whose presence 
or absence is given by the binary variable $D_{i,t}$, the details of this model 
are in the next section.

@@

\subsection{Parametric bootstrap for risk-reduction evaluation problems}
\label{varbootstrap}

In this section, we present a parametric bootstrap procedure for our AR(1) model (see Section \ref{modelsetup}) in approximating the distribution of our shock-effect estimators. %NO NEED TO OUTLINE IN THIS WAY: In the first step, we will introduce the pros and cons of this algorithm and refer the users to other alternatives. In the second step, we detail the procedure of boostrapping AR(1).
%%\citet{efron} introduces the bootstrapping method to approximate the distribution of a statistic using a \emph{random sample}. However, this method will fail for time-series data since serial correlation or time-dependent features exist in typical time-series model. 
In the AR($p$) model, the unobserved errors are assumed to be identically and independently distributed. In this setting, standard bootstrap methodology can be applied to resampling the residuals \citep{efron1986bootstrap}. \citet{bose1988edgeworth} showed that the asymptotic accuracy for OLS parameter estimation can be further improved from $O(T^{-1/2})$ to $o(T^{-1/2})$ almost surely under some regularity conditions. Nevertheless, the pseudo time series generated by this procedure are not stationary.

\citet{politis1994stationary} motivated a stationary bootstrap method for strictly stationary and weakly dependent time series $\{X_n \colon n \in \mathbb{N}\}$. This algorithm generates a sequence of blocks of observations $B_{I_1, L_1}, B_{I_2, L_2}, \ldots$, where $B_{i,b}= \{X_i, X_{i+1}, \ldots, X_{i+b-1}\}$\footnote{for $j > N$, $X_{j}$ is defined to be $X_k$, where $k = j \hspace{-.1cm} \mod N$ and $X_0 = X_N$}; and the sampling stops when $N$ observations are reached. Note that the collection of random positions $\{I_n \colon n \in \mathbb{N}\}$ is a sequence of i.i.d. discrete uniform random variables; and the collection of random lengths $\{L_n \colon n \in \mathbb{N}\}$ is a sequence of  i.i.d. geometric random variables with parameter $p$. However, the consistency needs to be proved by a case-by-case analysis \citep[Page 66]{politis1999subsampling}. Additionally, the asymptotic accuracy of this algorithm can be sensitive to the selection of $p$. This issue is similar to that of the selection of block size in moving-block bootstrapping \citep{kunsch1989jackknife, liu1992moving}.  More work related to bootstrapping time series  can be referred to Chapters 3 and 4 in \citet{politis1999subsampling}, and \citet{berkowitz2000recent}. It is up to the user in selecting which procedure to choose but under different assumptions on the time-series.

\vspace{.2cm}
\begin{algorithm}[H]
\SetAlgoLined \label{boots}

\SetKwInOut{Input}{Input}
\Input{$B$ --  the number of parametric bootstraps  \\  $\{(y_{i,t}, \mbf{x}_{i,t})\colon i = 2,\ldots, n+1, t = 0, \ldots, T_i\}$ -- the data\\ $\{T_i^* \colon i = 1, \ldots, n+1\}$ -- the time point just before the shock\\$\{\hat{\varepsilon}_{i, t} \colon t = 1, \ldots, T_i\}$ -- the collection of  residuals for $t = 1,\ldots, T_i$ \\ $\{\hat{\eta}_i, \hat{\alpha}_i, \hat{\phi}_i, \hat{\theta}_i, \hat{\beta}_i\colon i = 2, \ldots, n+1\}$ -- the OLS estimates \\
}

\KwResult{The sample mean, and sample variance of bootstrapped adjustment estimator, inverse-variance weighted estimator, and weighted-adjustment estimator.}

  \For{$b = 1:B$}{
  \For{$i = 2, \ldots, n+1$}{

    Sample with replacement from $\{\hat{\varepsilon}_{i, t} \colon t = 1, \ldots, T_i\}$ to obtain $\{\hat{\varepsilon}_{i, t}^{(b)} \colon t = 1, \ldots, T_i\}$

     Define $y_{i,0}^{(b)}=y_{i,0}$

    \For{$t = 1,\ldots, T_i$}{

     Compute $y_{i,t}^{(b)}=\hat{\eta}_i +\hat{\alpha}_i 1(t=T_i^*+1)+\hat{\phi}_iy_{i,t-1}^{(b)}+\theta_i'\mbf{x}_{i,t} + \beta_i'\mbf{x}_{i, t-1}+\hat{\varepsilon}_{i,t}^{(b)}$
    }
    Compute $\hat{\alpha}_i^{(b)}$ based on OLS estimation of (\ref{equation1}) with $\{(y_{i,t}^{(b)}, \mbf{x}_{i,t}) \colon t= 0, \ldots, T_i\}$
  }
   Compute the $b$th shock-effect estimate $\hat{\alpha}^{(b)}_{\rm est}$ for $\mrm{est} \in \{\mrm{adj}, \mrm{wadj}, \mrm{IVW}\}$
  }
   Compute the sample mean, and sample variance of $\{\hat{\alpha}_{\rm est}^{(b)} \colon b = 1, \ldots, B\}$ for $\mrm{est} \in \{\mrm{adj}, \mrm{wadj}, \mrm{IVW}\}$
 \caption{Parametric bootstrap for approximation for  mean and  variance of shock-effect estimators of $\alpha_1$.}
\end{algorithm}
\vspace{.2cm}

The procedures of bootstrapping AR(1) can be outline as in Algorithm \ref{boots}. We stress that the $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$ for  $t \in \{1, \ldots, T_i^*\}$ and $t' \in \{T_i^*+1, \ldots, T_i\}$ are i.i.d. since $y_{i,t}$ differs from $y_{i,t'}$ only in the shock-effect $\alpha_i$ conditioned on $\Theta$ (see Section \ref{modelsetup}), and it is \emph{not} absorbed into $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$. Therefore, this parametric bootstrap framework will work for us.

The estimates yielded by Algorithm \ref{boots} provides an approximation for parameters involved in the risk-reduction conditions in Sections \ref{conditions} and \ref{comparisons}. The almost-sure convergence results for $\hat{\alpha}_{\mrm{adj}}$  and $\hat{\alpha}_{\mrm{wadj}}$ follow naturally if we assume the regularity conditions outlined in  \citet{bose1988edgeworth}.
The proof for the case of $\overline{\hat{\alpha}_{\mrm{wadj}}}$, the sample mean of the bootstrapped weighted adjustment estimator, can be presented as below. As $B \to \infty$,
\begin{align*}
  \overline{\hat{\alpha}_{\mrm{wadj}}} \mid \Theta 
  = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)} \mid \Theta 
  = \sum_{i=2}^{n+1}w_i^* \frac{1}{B} \sum_{b=1}^B\hat{\alpha}_i^{(b)} \mid \Theta 
  \overset{a.s.}{\rightarrow} \sum_{i=2}^{n+1}w_i^* \alpha_i \mid \Theta=\E{\hat{\alpha}_{\rm wadj}|\Theta},
\end{align*}
The case for the sample variance of the bootstrapped weighted adjustment estimator would be similar. The same line of analysis holds for   adjustment estimator as well since  it is a linear combination of OLS estimates, and the weights are not random conditioned on $\Theta$ (see Section \ref{modelsetup}). Nevertheless, the case for $\hat{\alpha}_{\rm IVW}$ is slightly different because it is not a linear combination of OLS estimates with random weights. It is not clear with respect to whether similar consistency results holds for $\hat{\alpha}_{\rm IVW}$. We would claim that Algorithm \ref{boots} only provides an \emph{approximation} for the case of $\hat{\alpha}_{\rm IVW}$.

In particular $\overline{\hat{\alpha}_{\rm wadj}}$, the sample mean of $\hat{\alpha}_{\rm wadj}^{(b)}$, will provide an approximation for $\E{\alpha_1}$ under $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ since it is unbiased for $\E{\alpha_1}$ under those three configurations from Proposition \ref{unbiased}. Therefore, we can judge whether to use a shock-effect estimator, and choose between  shock-effect estimators by this method.

We \emph{stress} that the above approximation is conditioned on $\Theta$. Under a random-effect structure (see Section \ref{modelsetup}), every true shock effect of a time series in reality is only an observed value from a population.
  For example, $\overline{\hat{\alpha}_{\rm wadj}}\overset{a.s.}{\rightarrow}\E{\hat{\alpha}_{\rm wadj}|\Theta}$ and $\overline{\hat{\alpha}_{\rm wadj}} \not\overset{a.s.}{\rightarrow}\E{\alpha_1}$ but $\E{\E{\hat{\alpha}_{\rm wadj}|\Theta}}=\E{\alpha_1}$.

The simulation for justification of the parametric bootstrap is provided in Section \ref{parametricbootstrapsimulation}.

@@

In this section, we present the simulation for the parametric bootstrap procedure for AR(1) outlined in Section \ref{varbootstrap}. The simulation is only provided for the adjustment estimator and weighted adjustment estimator since the weights for inverse-variance estimator are random and the true values of its moments are unknown.

\begin{center}
  \begin{tabular}{l|cccccc}
    & \multicolumn{6}{c}{Time series}\\
    & 1 & 2 & 3 & 4 & 5 & 6\\
   Length $T_i$ & 77 & 93 & 72 & 71 & 73 & 50 \\
   Shock Time  $T_i^*+1$ & 20 & 39 & 45 & 52 & 59 & 11 \\
   Autoregressive parameter $\phi_i$ & 0.540 & 0.961 & 0.654 & 0.547 & 0.266 & 0.197 
  \end{tabular}
\end{center}

\section{Example}

In this example we forecast Conoco Phillips stock prices in the midst of the coronavirus (COVID-19) recession. Specific interest is in predictions made after March 6th, 2020, the Friday before the stock market crash on March 9th, 2020. We will detail how we  combine knowledge from disparate time series to improve the forecast of Conoco Phillips stock price without adjustment for the shock. The forecast consists of four steps: (1) pick a model, (2) selection of covariates,  (3) choices of donor pool, and (4) nowcast. Section \ref{background} provides the justification for the first three steps. Section \ref{results} discusses the result of our forecast.

\subsection{Background}
\label{background}
 Conoco Phillips is chosen for this analysis because it is a large oil and gas resources company \citep{conocowhatwedo} and with a relatively recent initial public offering \citep{conocohist}. Focus on the oil sector is because oil prices have been shown to exhibit a cointegrating behavior with economic indices \citep{he2010global}, and our chosen time frame represents the onset of a significant economic down turn, coupled with OPEC decreasing the global price of oil the Sunday before trading resumes on Monday, March 9th \citep{cnbc}. Furthermore, fear of and action in response to the coronavirus pandemic began to uptick dramatically between Friday, March 6th and Monday, March 9th. Major events include the SXSW festival being cancelled as trading closed on March 6th \citep{cnn}. New York declared a state of emergency on March 7th \citep{nygov}, and by Sunday, March 8th, eight states have declared a state of emergency while Italy placed 16 million people in quarantine \citep{newsweek}.

Economic indicators forecasted our recession before the coronavirus pandemic began. The current recession followed an inversion of the yield curve that first happened back in March 22, 2019 \citep{ap}. An inversion of the yield curve is an event that signals that recessions are more likely \citep{fed, frbsf, tokic2019yield}. In this analysis we investigate the performance of oil companies in previous recessions that followed an inversion of the yield curve to obtain a suitable Conoco Phillips donor pool for estimating the March 9th shock effect on Conoco Phillips oil stock. We also consider previous OPEC oil supply shocks. We will borrow from the literature on oil price forecasting to establish appropriate time horizons and forecasting models. Recessions that occurred before 1973 are disregarded since oil price forecasts cannot be represented by standard time series models before 1973 \citep{alquist2013forecasting}.

@@

 We will use a simple AR(1) model to forecast Conoco Phillips stock price. This model has been shown to beat no-change forecasts when predicting oil prices over time horizons of 1 and 3 months \citep{alquist2013forecasting}. We will consider 30 pre-shock trading days and we will forecast the immediate shock effect and the shock effect over a future five trading day window. All estimates will be adjusted for inflation using the inflation calculator at \url{https://www.usinflationcalculator.com}. The model setup for AR(1) is exactly the same as what is stated in Section \ref{modelsetup} with addition of shock effects. All the parameters are estimated using OLS.

@@

We perform two lines of analyses with one incorporating the price of S\&P 500 index and one without to justify our choice of covariate. Since the stock price of a company implies its value provided the number of shares is unchanged, the price of S\&P 500 index in the recession can partially represent the overall status of the economy.  Therefore, using the price of S\&P 500 index as the covariate can possibly improve the precision of the OLS estimates for shock-effects that are unique to Conoco Phillips. The rationale for inclusion of the price of S\&P 500 index is similar to that of market return in capital asset pricing model (CAPM).

@@

Our donor pool will consist of Conoco Phillips shock effects observed on September 11th, 2001, September 8th, 2008, and November 27, 2014. The first two shock effects were observed during recessions that were predicated by an inversion of the yield curve, and the third was an OPEC induced supply side shock effect. The reasons for those three shocks are
\begin{enumerate}
  \item On September 11th, 2001, Islamic extremist al-Qaeda  committed a series of four terrorist attacks  against U.S. It triggers negative percussions on the U.S. economy with falling stock prices, increased unemployment rates, and falling U.S. dollar \citep{nytimes911, history}.
  \item On September 8th, 2008, the United States announced its takeover of Fannie May and Freddie Mac \citep{UST}, and it was reported that the nation’s unemployment rate reached five-year highs. It is also worth noting that Lehman Brothers Holdings filed for bankruptcy on Monday, September 15th, 2008 with what was a record of 613 billion dollars in debt \citep{lehman}.
  \item On November 27th, 2014, it is documented that oil prices fall as OPEC opts not to cut production \citep{CNBC2014}. During the Great Recession when economic activity clearly declined, both oil and stock prices fell which points to demand factors. During the second half of 2014, oil prices plummeted but equity prices generally increased, suggesting that supply factors were the key driver \citep[Page 19]{baffes2015great}. 
\end{enumerate}

Some other notable shock effects were not included in this analysis because the conditions preceding them are not similar to those of the March 9, 2020 shock effect, among these are the October 19, 1987, August 8th, 2011, and  August 24, 2015 shock effects \citep{shiller1988portfolio, latimes, nytimes}.

@@

\section{Supplementary Materials}

@@

\begin{proof-of-proposition}[\ref{propriskwadj2}]
  By Proposition \ref{unbiased}, the property that $\hat{\alpha}_{\rm wadj}$ is an unbiased estimator of $\E{\alpha_1}$ holds for $\mc{M}_{21}$ and $\mc{M}_{22}$. The remainder of the proof follows a similar argument to the proof of Proposition \ref{proprisk} (\textbf{Fix this problematic reference}).
\end{proof-of-proposition}

>>>>

We will suppose that an analyst has time series data ($y_{i,t}$,$\x_{i,t}$), 
$t = 1$, $\ldots$, $T_i$, $i = 1$, $\ldots$, $n+1$, where $y_{i,t}$ is a 
scalar response and $\x_{i,t}$ is a vector of covariates that are revealed to 
the analyst prior to the observation of $y_{1,t}$.  Suppose that the analyst 
is interested in forecasting $y_{1,t}$, the first time series in the 
collection.
%$y_1, y_2, \ldots$.  
%Given each time point $t \geq 1$, let $\x_{1,t}$ be a vector of covariates 
%revealed prior to the observation of $y_{1,t}$.  
To gauge the performance of a procedure that produces forecasts 
$\{\hat y_{1,t}, t= 1,2,\ldots\}$ given time horizon $T_1$, 
we consider forecast risk in the form of root mean squared error
$$
  R_T = \sqrt{\frac{1}{T}\sum_{t=1}^T\E{\hat y_{1,t} - y_{1,t}}^2}
$$
in our analyses. \textbf{Let's switch to root mean squared error, this metric is often used in forecasting articles.} In this article, we consider a dynamic panel data 
model with autoregressive structure similar to that in \citet{blundell1998initial}. 
%without past time period covariate 
%information and with a single present time covariate.  
Our dynamic panel model includes an additional shock effect whose presence 
or absence is given by the binary variable $D_{i,t}$, the details of this model 
are in the next section.

@@

\subsection{Bootstrap for risk-reduction evaluation problems}
\label{varbootstrap}

In this section, we present a bootstrap procedure 
%(see Section \ref{modelsetup}) 
that approximates the distribution of our shock-effect estimators and check the underlying conditions of our risk reduction results in practice. Our procedure involves the resampling of residuals in the separate OLS fits. This procedure has its origins in Section 6 of \citet{efron1986bootstrap}, and it involves the resampling the residuals which are assumed to be the realizations of an iid processes.
%NO NEED TO OUTLINE IN THIS WAY: In the first step, we will introduce the pros and cons of this algorithm and refer the users to other alternatives. In the second step, we detail the procedure of boostrapping AR(1).
%%\citet{efron} introduces the bootstrapping method to approximate the distribution of a statistic using a \emph{random sample}. However, this method will fail for time-series data since serial correlation or time-dependent features exist in typical time-series model. 
%In the AR($p$) model, the unobserved errors are assumed to be identically and independently distributed. In this setting, standard bootstrap methodology can be applied to resampling the residuals \citep{efron1986bootstrap}. 
\citet{bose1988edgeworth} showed that the asymptotic accuracy for OLS parameter estimation can be further improved from $O(T^{-1/2})$ to $o(T^{-1/2})$ almost surely under some regularity conditions. However, the pseudo time series generated by this procedure are not stationary. (\textbf{Is this necessary to state? There needs to be a rationale fro why this is not a problem.})

\textbf{I think that our bootstrap procedure should first sample the disparate time series without replacement, and then estimate the parameters in each resampled time series.}
The formal steps of our bootstrap procedures are outlined in the Supplementary Materials, the intuition for our procedure is as follows: let $B$ be the bootstrap sample size and initialize $y_{i,0}$ for all $i = 2,\ldots,n+1$. At iteration $b$, resample the residuals and then obtain shock-effect estimators for each of the disparate time series for all $i = 2,\ldots,n+1$. Then construct and store any of the adjustment estimators $\hat{\alpha}^{(b)}_{\mrm{adj}}$, $\hat{\alpha}^{(b)}_{\mrm{wadj}}$, and $\hat{\alpha}^{(b)}_{\mrm{IVW}}$. We can then estimate distributional quantities of our shock-effect estimators with the bootstrap sample of $\hat{\alpha}^{(b)}_{\mrm{adj}}$, $\hat{\alpha}^{(b)}_{\mrm{wadj}}$, and $\hat{\alpha}^{(b)}_{\mrm{IVW}}$, for $b = 1,\ldots,B$.
%The errors $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$ for $t \in \{1, \ldots, T_i^*\}$ and $t' \in \{T_i^*+1, \ldots, T_i\}$ are iid. 
%since $y_{i,t}$ differs from $y_{i,t'}$ only in the shock-effect $\alpha_i$ conditioned on $\Theta$, and it is not absorbed into $\varepsilon_{i,t}$ and $\varepsilon_{i,t'}$. Therefore, this parametric bootstrap framework will work for us.
Once can then use the bootstrapped shock-effects estimated by our residual bootstrap to provide an approximation for parameters involved in the risk-reduction conditions in Propositions \ref{proprisk} and \ref{propriskwadj2}.

Let 
$
  \overline{\hat{\alpha}_{\mrm{aadj}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}\hat{\alpha}_i^{(b)}
$, 
$
  \overline{\hat{\alpha}_{\mrm{wadj}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)}
$, 
and 
$
  \overline{\hat{\alpha}_{\mrm{IVW}}} 
    = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}\hat{\alpha}_{{\rm IVW}, i}^{(b)}.
$
The almost-sure convergence results for $\hat{\alpha}_{\mrm{adj}}$ and $\hat{\alpha}_{\mrm{wadj}}$ follow naturally, provided that the regularity conditions outlined in \citet{bose1988edgeworth} hold. The proof for the case of $\overline{\hat{\alpha}_{\mrm{wadj}}}$, the sample mean of the bootstrapped weighted adjustment estimator, can be presented as below. As $B \to \infty$,
\begin{align*}
  \overline{\hat{\alpha}_{\mrm{wadj}}}
  = \frac{1}{B} \sum_{b=1}^B \sum_{i=2}^{n+1}w_i^* \hat{\alpha}_i^{(b)}
  = \sum_{i=2}^{n+1}w_i^* \frac{1}{B} \sum_{b=1}^B\hat{\alpha}_i^{(b)}
  \overset{a.s.}{\rightarrow} \sum_{i=2}^{n+1}w_i^* \alpha_i 
  =\E{\hat{\alpha}_{\rm wadj}|\Theta},
\end{align*}
The case for the sample variance of the bootstrapped weighted adjustment estimator is similar. The same rationale holds for the adjustment estimator as well since it is a linear combination of OLS estimates, and the weights are not random conditioned on $\Theta$. However, the case for $\hat{\alpha}_{\rm IVW}$ is slightly different because it is a linear combination of OLS estimates with random weights. It is not clear with respect to whether similar consistency results holds for $\hat{\alpha}_{\rm IVW}$. We claim that our bootstrap procedure provides an approximation for the case of $\hat{\alpha}_{\rm IVW}$.

%In particular $\overline{\hat{\alpha}_{\rm wadj}}$ will provide an approximation for $\E{\alpha_1}$ under $\mc{M}_1$, $\mc{M}_{21}$, and $\mc{M}_{22}$ since it is unbiased for $\E{\alpha_1}$ under those three configurations from Proposition \ref{unbiased}. Therefore, we can judge whether to use a shock-effect estimator, and choose between shock-effect estimators by this method.

We stress that the above approximation is conditioned on $\Theta$ and that bootstrapping cannot alleviate the inherent bias of using our adjustment estimators as estimates for $\alpha_1$. %Under a random-effect structure, every true shock effect of a time series in reality isan observed value from a population. For example, $\overline{\hat{\alpha}_{\rm wadj}}\overset{a.s.}{\rightarrow}\E{\hat{\alpha}_{\rm wadj}|\Theta}$ and $\overline{\hat{\alpha}_{\rm wadj}} \not\overset{a.s.}{\rightarrow}\E{\alpha_1}$ but $\E{\E{\hat{\alpha}_{\rm wadj}|\Theta}}=\E{\alpha_1}$. 
Simulation for justification of the parametric bootstrap is provided in Section \ref{parametricbootstrapsimulation}.

@@

%In this section, we present the simulation for the parametric bootstrap procedure for AR(1) outlined in Section \ref{varbootstrap}. The simulation is only provided for the adjustment estimator and weighted adjustment estimator since the weights for inverse-variance estimator are random and the true values of its moments are unknown.

%\begin{center}
%  \begin{tabular}{l|cccccc}
%    & \multicolumn{6}{c}{Time series}\\
%    & 1 & 2 & 3 & 4 & 5 & 6\\
%   Length $T_i$ & 77 & 93 & 72 & 71 & 73 & 50 \\
%   Shock Time  $T_i^*+1$ & 20 & 39 & 45 & 52 & 59 & 11 \\
%   Autoregressive parameter $\phi_i$ & 0.540 & 0.961 & 0.654 & 0.547 & 0.266 & 0.197 
%  \end{tabular}
%\end{center}

\section{Forecasting Conoco Phillips stock in the presence of shocks}

In this example we forecast Conoco Phillips stock prices in the midst of the coronavirus (COVID-19) recession. Specific interest is in predictions made after March 6th, 2020, the Friday before the stock market crash on March 9th, 2020. We will detail how we combine knowledge from disparate time series to improve the forecast of Conoco Phillips stock price that would be made without adjustment for the shock. The forecast consists of four steps: (1) pick a model, (2) selection of covariates,  (3) choices of donor pool, and (4) nowcast. Section \ref{background} provides the justification for the first three steps. Section \ref{results} discusses the result of our forecast.

\subsection{Background}
\label{background}

 Conoco Phillips is chosen for this analysis because it is a large oil and gas resources company \citep{conocowhatwedo}. %and with a relatively recent initial public offering \citep{conocohist}. 
Focus on the oil sector is because oil prices have been shown to exhibit a cointegrating behavior with economic indices \citep{he2010global}, and our chosen time frame represents the onset of a significant economic down turn, coupled with a Russia and OPEC battle for global oil price control the Sunday before trading resumes on Monday, March 9th \citep{sukhankin2020russian}. Furthermore, fear of and action in response to the coronavirus pandemic began to uptick dramatically between Friday, March 6th and Monday, March 9th. Major events include the SXSW festival being cancelled as trading closed on March 6th \citep{wang2020impact}. New York declared a state of emergency on March 7th \citep{nygov}, and by Sunday, March 8th, eight states have declared a state of emergency \citep{alonso2020state} while Italy placed 16 million people in quarantine \citep{sjodin2020only}.

Economic indicators forecasted our recession before the coronavirus pandemic began. The current recession followed an inversion of the yield curve that first happened back in March, 2019 \citep{tokic2019yield}. An inversion of the yield curve is an event that signals that recessions are more likely \citep{andolfatto2018yield, bauer2018economic}. In this analysis we investigate the performance of oil companies in previous recessions that followed an inversion of the yield curve to obtain a suitable Conoco Phillips donor pool for estimating the March 9th shock effect on Conoco Phillips oil stock. We also consider previous OPEC oil supply shocks. We will borrow from the literature on oil price forecasting to establish appropriate time horizons and forecasting models. Recessions that occurred before 1973 are disregarded since oil price forecasts cannot be represented by standard time series models before 1973 \citep{alquist2013forecasting}.

@@

 We will use a simple AR(1) model to forecast Conoco Phillips stock price. This model has been shown to beat no-change forecasts when predicting oil prices over time horizons of 1 and 3 months \citep{alquist2013forecasting}. We will consider 30 pre-shock trading days and we will forecast the immediate shock effect and the shock effect over a future five trading day window. All estimates will be adjusted for inflation. The model setup for AR(1) is exactly the same as what is stated in Section \ref{modelsetup} with addition of shock effects. All the parameters are estimated using OLS.

@@

We perform two lines of analyses with one incorporating the price of S\&P 500 index and one without to justify our choice of covariate. Since the stock price of a company implies its value provided the number of shares is unchanged, the price of S\&P 500 index in the recession can partially represent the overall status of the economy.  Therefore, using the price of S\&P 500 index as the covariate can possibly improve the precision of the OLS estimates for shock-effects that are unique to Conoco Phillips. The rationale for inclusion of the price of S\&P 500 index is similar to that of market return in capital asset pricing model (CAPM). (\textbf{Need reference for the CAPM}.)

@@

Our donor pool will consist of Conoco Phillips shock effects observed after September 11th, 2001, %March 18th, 2008, 
several events in September, 2008, and November 27, 2014. The first two shock effects were observed during recessions that were predicated by an inversion of the yield curve \citep{bauer2018economic}, and the third was an OPEC induced supply side shock effect \citep{huppmann2015opec}. The reasons for those three shocks are
\begin{enumerate}
  \item On September 11th, 2001, Islamic extremist al-Qaeda committed a series of terrorist attacks against the United States of America \citep{braniff2011towards}. In addition to the tragic loss of life, this triggered negative repercussions on the U.S. and world economies with falling stock prices and a falling U.S. dollar \citep{floyd2001terror}. Moreover, trading was closed until September 17, 2001 \citep{johnston2006impact}.
  %\item In heavy trading on Tuesday, March 18 \citep{shorter2008bear}
  \item On September 8th, 2008, the United States announced its takeover of Fannie May and Freddie Mac \citep{UST}, and it was reported that the nation’s unemployment rate reached five-year highs. It is also worth noting that Lehman Brothers Holdings filed for bankruptcy on Monday, September 15th, 2008 with what was a record of 613 billion dollars in debt \citep{lehman}.
  \item On November 27th, 2014, it is documented that oil prices fall as OPEC opts not to cut production \citep{huppmann2015opec}. During the Great Recession when economic activity clearly declined, both oil and stock prices fell which points to demand factors. During the second half of 2014, oil prices plummeted but equity prices generally increased, suggesting that supply factors were the key driver \citep[Page 19]{baffes2015great}. 
\end{enumerate}

%Some other notable shock effects were not included in this analysis because the conditions preceding them are not similar to those of the March 9, 2020 shock effect, among these are the October 19, 1987, August 8th, 2011, and  August 24, 2015 shock effects \citep{shiller1988portfolio, latimes, nytimes}.

@@

\section{Discussion}

Our bootstrap procedure can be extended to approximate the distribution of shock effect estimators from more general time series. \citet{politis1994stationary} motivated a stationary bootstrap method for strictly stationary and weakly dependent time series $\{X_n \colon n \in \mathbb{N}\}$. This algorithm generates a sequence of blocks of observations $B_{I_1, L_1}, B_{I_2, L_2}, \ldots$, where $B_{i,b}= \{X_i, X_{i+1}, \ldots, X_{i+b-1}\}$\footnote{for $j > N$, $X_{j}$ is defined to be $X_k$, where $k = j \hspace{-.1cm} \mod N$ and $X_0 = X_N$}; and the sampling stops when $N$ observations are reached. Note that the collection of random positions $\{I_n \colon n \in \mathbb{N}\}$ is a sequence of i.i.d. discrete uniform random variables; and the collection of random lengths $\{L_n \colon n \in \mathbb{N}\}$ is a sequence of  i.i.d. geometric random variables with parameter $p$. However, the consistency needs to be proved by a case-by-case analysis \citep[Page 66]{politis1999subsampling}. Additionally, the asymptotic accuracy of this algorithm can be sensitive to the selection of $p$. This issue is similar to that of the selection of block size in moving-block bootstrapping \citep{kunsch1989jackknife, liu1992moving}.  More work related to bootstrapping time series  can be referred to Chapters 3 and 4 in \citet{politis1999subsampling}, and \citet{berkowitz2000recent}. It is up to the user in selecting which procedure to choose but under different assumptions on the time-series.

\section{Appendix}

@@

\begin{proof-of-proposition}[\ref{propriskwadj2}]
  By Proposition \ref{unbiased}, the property that $\hat{\alpha}_{\rm wadj}$ is an unbiased estimator of $\E{\alpha_1}$ holds for $\mc{M}_{21}$ and $\mc{M}_{22}$. The remainder of the proof follows a similar argument to the proof of Proposition \ref{proprisk}.
\end{proof-of-proposition}
